Daily goals

### 2020-06-24 docs/submission-mee-2020-06-24.tex
mee

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Obtention of molecular or morphological character data -- get DNA from some organisms
  and sequence it, or get it from an online nucleotide data repository, such as GenBank (Benson et al. \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler et al. \protect\hyperlink{ref-wheeler2000database}{2000}).
\item
  Assemble a hypothesis of homology -- Create a matrix of your character data, by
  aligning the sequences, in the case of molecular data. Make sure thay are paralogs!
\item
  Analyse this hypothesis of homology to infer phylogenetic relationships among
  the organisms you are studying -- Use different available programs to infer molecular
  evolution, trees and times of divergence.
\item
  Discuss the inferred relationships in the context of previous hypothesis, the
  biology and biogeography of the organisms, etc. -- Answer the question, \emph{is this phylogenetic solution fair/reasonable?}
\end{enumerate}

@@

In the past decade, various studies have developed solutions to automatize the first and second steps, by creating
pipelines that mine already available molecular data from the GenBank repository (Benson et al. \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler et al. \protect\hyperlink{ref-wheeler2000database}{2000}),
to obtain homologous characters that can be used for phylogenetic reconstruction.
These tools have been presented as aid for the nonspecialist to decrease some of
the difficulties in the generation of phylogenetic knowledge. However, they are
not that often used as so, suggesting that there are still difficulties for the nonspecialist.
The phylogenetic community has some reserves towards these tools, too. Mainly because they sometimes act as a black box.
However, automatizing the assembly of the character data set is a crucial step towards
reproducibility for a task that was otherwise primarily artisanal and hence largely
non-reproducible.

@@

Public phylogenies can be updated with the ever increasing amount of genetic data that is available on GenBank (Benson et al. \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler et al. \protect\hyperlink{ref-wheeler2000database}{2000}).

@@

The goal of Physcraper is to build upon previous phylogenetic knowledge,
allowing a direct comparison between existing phylogenies and phylogenies that are constructed
using new genetic data retrieved from a public nucleotide database (i.e., GenBank (Benson et al. \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler et al. \protect\hyperlink{ref-wheeler2000database}{2000})).

@@

\begin{itemize}
\tightlist
\item
  The study tree is a published phylogenetic tree stored in the OToL database, phylesystem (McTavish et al. \protect\hyperlink{ref-mctavish2015phylesystem}{2015}). The main
  reason for this is that trees in phylesystem have a set of user defined characteristics
  that are essential for automatizing the phylogeny update process. The most relevant of these being the definition of ingroup and outgroup. Outgroup and ingroup taxa in the original tree are identified and tagged. This allows to automatically set the root for the updated tree on the next steps of the pipeline.
  A user can choose from the `r rotl::tol\_about()\$num\_source\_trees' published trees supporting the resolved node of the synthetic tree in the OToL website (\textless{}\textgreater{}). If the tree you are interested in updating is not in there, you can upload it via OToL's curator tool (\textless{}\url{https://tree.opentreeoflife.org/curator}).
\item
  The alignment should be a gene alignment that was used to generate the tree. The original
  alignments are usually stored in a public repository such as TreeBase (Piel et al. \protect\hyperlink{ref-piel2009treebase}{2009}; Vos et al. \protect\hyperlink{ref-vos2012nexml}{2012}),
  DRYAD (\url{http://datadryad.org/}), or the journal were the tree was originally published.
  If the alignment is stored in TreeBase, \texttt{physcraper} can download it directly,
  either from the TreeBASE website (\url{https://treebase.org/})
  or through the TreeBASE GitHub repository (SuperTreeBASE; \url{https://github.com/TreeBASE/supertreebase}).
  If the alignment is on another repository, or provided personally by the owner, a copy of it has to be
  downloaded by the user, and it's local path has to be provided as an argument.
\item
  A taxon name matching step is performed to verify that all taxon names on the tips
  of the tree are in the DNA character matrix and vice versa.
\item
  A ``.csv'' file with the summary of taxon name matching is produced for the user.
\item
  Unmatched taxon names are dropped from both the tree and alignment.
  Technically, just one matching name is needed to perform the searches. Please, see next section.
\item
  A ``.tre'' file and a ``.fas'' file containing only the matched taxa are generated and saved in the \texttt{inputs} folder to be used in the following steps.
\end{itemize}

@@

  Users can provide a search taxon that is either a more or a less inclusive
  clade relative to the ingroup of the original phylogeny. If the search taxon is more inclusive, the sequence search will be performed outside the MRCAT of the matched taxa, e.g., including all taxa within
  the family or the order that the ingroup belongs to. If the search taxon is a less inclusive clade, the users can focus on enriching a particular clade/region within the ingroup of the phylogeny.
\item
  The Basic Local Alignment Search Tool, BLAST {[}Altschul et al. (\protect\hyperlink{ref-altschul1990basic}{1990}); altschul1997gapped{]} is used to identify
  similarity between DNA sequences within the search taxon in a nucleotide
  database, and the remaining sequences on the alignment.
  The BLAST command line tools (Camacho et al. \protect\hyperlink{ref-camacho2009blast}{2009}) are used for both web-based and local-database searches.
\item
  A pairwise alignment-against-all BLAST search is performed. This means that each sequence
  in the alignment is BLASTed against DNA sequences in a nucleotide database constrained to the search
  taxon. Results from each one of these BLAST runs are recorded, and matched sequences are saved
  along with their corresponding identification numbers (accesion numbers in the case of the GenBank database). This information will be used later to store the whole sequences in a dedicated library within the physcraper folder, allowing for secondary analyses to run significantly faster.
\item
  The DNA sequence similarity search can be done on a local database that is easily
  setup by the user. In this case, the BLASTn algorithm is used to performs the similarity search.
\item
  The search can also be performed remotely, on the NCBI database. In this case, the
  bioPython BLAST algorithm is used to perform the similarity search.
\item
  Matched sequences below an e-value, percentage similarity, and outside a minimum
  and maximum length threshold are discarded. This filtering leaves out genomic sequences.
  All acepted sequences are asigned an internal identifier, and are further filtered.
\item
  Because the original alignments usually lack database id numbers, a filtering
  step is needed. Accepted sequences that belong to the
  same taxon of the query sequence, and that are either identical or shorter than
  the original sequence are discarded. Only longer sequences belonging to the
  same taxon as the orignal sequence will be considered further for analysis.
\item
  Among the remaining filtered sequences, there are usually several exemplars per taxon.
  Although it can be useful to keep some of them to, for example, investigate monophyly
  within species, there can be hundreds of exemplar sequences per taxon for some markers.
  To control the number of sequences per taxon in downstream analyses,
  5 sequences per taxon are chosen at random. This number is set by default but can be modified by the user.
\item
  Reverse complement sequences are identified and translated.
\item
  Users can choose to perform a more ``cycles'' of sequence similarity search, by blasting the newly found sequences. This can be done iteratively, but by default only sequences in the alignment are blasted. \textbf{\emph{Is there an argument to control the number of cycles of blast searches with new sequences?}}
\item
  Accepted sequences are downloaded in full, and stored as a local database in a directory that is globally accesible (physcraper/taxonomy), so they are accesible for further runs.
\item
  A fasta file containing all filtered and processed sequences resulting from the BLAST search is generated for the user.
\end{itemize}

@@

\begin{itemize}
\tightlist
\item
  A gene tree is reconstructed for each alignment provided, using the software RAxML (Stamatakis \protect\hyperlink{ref-stamatakis2014raxml}{2014})
  with 100 classic bootstrap (Felsenstein \protect\hyperlink{ref-felsenstein1985confidence}{1985}) replicates by default. The number of bootsrap replicates can be modified by the user.
  Other type of bootstrap that I think is not yet incorporated into physcraper is the Transfer Bootstrap Expectation (TBE) recently proposed in Lemoine et al. (\protect\hyperlink{ref-lemoine2018renewing}{2018}).
\item
  The final result is an updated phylogenetic hypothesis for each of the genes provided in the alignment.
\item
  Tips on all trees generated by physcraper are defined by a taxon name space, allowing to perform comparisons and conflict analyses.
\item
  Robinson Foulds metrics
\item
  Describe what a conflict analysis is: Node by node comparison of the resulting clades compared to
\item
  For the conflict analysis to be meaningful, the root of the tree ineeds to be accurately defined.
\item
  Currently, the root is determined by finding the parent node of the sequences that do not belong to the ingroup/ search taxon. This ensures a correct rooting of the tree even when the search taxon is more inclusive than the ingroup.
\item
  Conflict information can only be generated in the context of the whole Open Tree of
  Life. Otherwise, it is not really possible to get conflict data.
  \textbf{\emph{- One way to compare two independent phylogenetic trees is to compare them both to
  the synthetic OToL and then measure how well they do against each other}}
\end{itemize}

@@

The genus \emph{Ilex} is the only extant clade within the family Aquifoliaceae, order Aquifoliales of flowering plants.
It encompasses between 400-600 living species. A review of litterature shows that there are three published phylogenetic trees, showing relationships within the hollies.
The first one has been made available both on OToL phylesystem and synth tree, and on treeBASE, it samples 48 species.
The second has not been made available anywhere, not even in supplementary data of the journal.
\textbf{\emph{Contact authors? They seem old school, probably do not wanna share their data.}}
The most recent one has been made available in the OToL Phylesystem and DRYAD. It is the best sampled yet, with 200 species. However,
it has not been added to the syntehtic tree yet.
This makes it a perfect case to test the basic functionalities of physcraper: we know that the sequences of the most recently published tree have been made available on the GenBank database (Benson et al. \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler et al. \protect\hyperlink{ref-wheeler2000database}{2000}). Updating the oldest tree, we should get something very similar to the newest tree.

@@

Let's be more specific now about our X group and say it is the Ascomycota.
The best tree currently available in OToL was published by Schoch et al. (\protect\hyperlink{ref-schoch2009ascomycota}{2009}).
The first step, is to get the Open Tree of Life study id. There are some options to do this:
- You can go to the Open Tree of Life website and browse until you find it, or
- you can get the study id using R tools:
- By using the TreeBase ID of the study (which is not fully exposed on the
TreeBase website home page of the study, so you have to really look it up manually):

@@

Phylogeny of the Testudines 6 tips from Crawford et al. (\protect\hyperlink{ref-crawford2012more}{2012})
There is just one tree in OToL.
There is just one alignment on \href{https://treebase.org/treebase-web/search/study/matrices.html?id=12742}{treebase} with all the 1 145 loci.

@@

Describe again statistics to compare phylogenies provided by physcraper via OpenTreeOfLife.
Mention statistics provided by other tools: PhyloExplorer (Ranwez et al. \protect\hyperlink{ref-ranwez2009phyloexplorer}{2009}).
Compare and discuss.

@@

How can it be used for the advantage of the field:
- rapid phylogenetic placing of newly discovered species, as mentioned in Webb et al. (\protect\hyperlink{ref-webb2010biodiversity}{2010})
- obtain trees for ecophylogenetic studies, as mentioned in Helmus and Ives (\protect\hyperlink{ref-helmus2012phylogenetic}{2012})
- one day could be used to sistematize nucleotide databases, such as Genbank (Benson et al. \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler et al. \protect\hyperlink{ref-wheeler2000database}{2000}), as mentioned in San Mauro and Agorreta (\protect\hyperlink{ref-san2010molecular}{2010}), i.e., curate ncbi taxonomic assignations.
- allows to generate custom species trees for downstream analyses, as mentioned in Stoltzfus et al. (\protect\hyperlink{ref-stoltzfus2013phylotastic}{2013})

Things that physcraper does not do:
- analyse the whole GenBank database (Benson et al. \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler et al. \protect\hyperlink{ref-wheeler2000database}{2000}) to find homolog regions suitable to reconstruct phylogenies, as mentioned in Antonelli et al. (\protect\hyperlink{ref-antonelli2017toward}{2017}). There are already some very good tools that do that.
- provide basic statistics on data availability to assemble molecular datasets, as mentioned by Ranwez et al. (\protect\hyperlink{ref-ranwez2009phyloexplorer}{2009}). Phyloexplorer does this?
- it is not a tree repo, as phylota is, mentioned in Deepak et al. (\protect\hyperlink{ref-deepak2014evominer}{2014})

@@

Sanderson et al. (\protect\hyperlink{ref-sanderson2008phylota}{2008})\strut

@@

Wu and Eisen (\protect\hyperlink{ref-wu2008simple}{2008})\strut

@@

Smith et al. (\protect\hyperlink{ref-smith2009mega}{2009})\strut

@@

Peters et al. (\protect\hyperlink{ref-peters2011taming}{2011})\strut

@@

Grant and Katz (\protect\hyperlink{ref-grant2014building}{2014})\strut

@@

Chesters and Zhu (\protect\hyperlink{ref-chesters2014protocol}{2014})\strut

@@

Izquierdo-Carrasco et al. (\protect\hyperlink{ref-izquierdo2014pumper}{2014})\strut

@@

Meng et al. (\protect\hyperlink{ref-meng2015darwintree}{2015}\protect\hyperlink{ref-meng2015darwintree}{a})\strut

@@

Xu et al. (\protect\hyperlink{ref-xu2015ncbiminer}{2015})\strut

@@

McMahon et al. (\protect\hyperlink{ref-mcmahon2015stbase}{2015})\strut

@@

Papadopoulou et al. (\protect\hyperlink{ref-papadopoulou2015automated}{2015})\strut

@@

Kumar et al. (\protect\hyperlink{ref-kumar2015bir}{2015})\strut

@@

Antonelli et al. (\protect\hyperlink{ref-antonelli2017toward}{2017})\strut

@@

Drori et al. (\protect\hyperlink{ref-drori2018onetwotree}{2018})\strut

@@

Smith and Walker (\protect\hyperlink{ref-smith2019pyphlawd}{2019})\strut

@@

Cerón-Romero et al. (\protect\hyperlink{ref-ceron2019phylotol}{2019})\strut

@@

Bennett et al. (\protect\hyperlink{ref-bennett2018phylotar}{2018})\strut

@@

According to Cerón-Romero et al. (\protect\hyperlink{ref-ceron2019phylotol}{2019}), PhyLoTA and BIR ``focus on the identification and collection
of homologous and paralog genes from public databases such as GenBank'', while both AMPHORA and PHLAWD
``focus on the construction and refinement of robust alignments rather than the collection of homologs.''

@@

PhyloFinder (Chen et al. \protect\hyperlink{ref-chen2008phylofinder}{2008}) - cited by 18: a search engine for phylogenetic databases, using
trees from TreeBASE - more related to phylotastic's goal than to updating/creating phylogenies

@@

PhyloExplorer (Ranwez et al. \protect\hyperlink{ref-ranwez2009phyloexplorer}{2009}) - cited by 21: a python and MySQL based website to facilitate
assessment and management of phylogenetic tree collections. It provides ``statistics describing the collection,
correcting invalid taxon names, extracting taxonomically relevant parts of the collection
using a dedicated query language, and identifying related trees in the TreeBASE database''.

@@

PhySpeTre (Fang et al. \protect\hyperlink{ref-fang2019physpetree}{2019}) - no citations yet - no sequence retrieval, just phylogenetic reconstruction
pipeline.

@@

Datataxa Ruiz-Sanchez et al. (\protect\hyperlink{ref-ruiz2019datataxa}{2019}) - no citations yet - focus on extracting metadata from GenBank sequence information.

@@

Phylota was published as a website to summarize and browse the phylogenetic potential of the GenBank
database (Sanderson et al. \protect\hyperlink{ref-sanderson2008phylota}{2008}).

@@

    original publication of PHLAWD (Smith et al. \protect\hyperlink{ref-smith2009mega}{2009})

@@

    the red algal tree of life (Verbruggen et al. \protect\hyperlink{ref-verbruggen2010data}{2010})

@@

    Beaulieu et al. (\protect\hyperlink{ref-beaulieu2012modeling}{2012}\protect\hyperlink{ref-beaulieu2012modeling}{a}) cites phylota as an example study of very large and comprehensive

@@

    a review for ecologists about phylogenetic tools (Roquet et al. \protect\hyperlink{ref-roquet2013building}{2013})

@@

    a study constructing a dated seed plant phylogeny using pyPHLAWD (Smith and Brown \protect\hyperlink{ref-smith2018constructing}{2018})

@@

    using genomic data. It aims to be incorporated into a pipeline such as phylota some day (Fan et al. \protect\hyperlink{ref-fan2015assembly}{2015}).

@@

    nexml format presentation (Vos et al. \protect\hyperlink{ref-vos2012nexml}{2012}) - cites phylota as a tool that uses

@@

    tips of the Campanulidae, constructed with PHLAWD (Beaulieu and Donoghue \protect\hyperlink{ref-beaulieu2013fruit}{2013})

@@

    a study of Southeast Asia plant biodiversity inventory (Webb et al. \protect\hyperlink{ref-webb2010biodiversity}{2010}) -

@@

    a study of wood density for carbon stock assessments (Flores and Coomes \protect\hyperlink{ref-flores2011estimating}{2011}),

@@

    a study proposing something similar to Open tree but applied only to land plants (Beaulieu et al. \protect\hyperlink{ref-beaulieu2012synthesizing}{2012}\protect\hyperlink{ref-beaulieu2012synthesizing}{b})

@@

    an analysis of the phylogenetic diversity-area curve (Helmus and Ives \protect\hyperlink{ref-helmus2012phylogenetic}{2012}),

@@

    (San Mauro and Agorreta \protect\hyperlink{ref-san2010molecular}{2010}) cites phylota as a tool ``intended to systematize GenBank information

>>>>

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Obtention of molecular or morphological character data -- get DNA from some organisms
  and sequence it, or get it from an online nucleotide data repository, such as GenBank (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000}).
\item
  Assemble a hypothesis of homology -- Create a matrix of your character data, by
  aligning the sequences, in the case of molecular data. Make sure thay are paralogs!
\item
  Analyse this hypothesis of homology to infer phylogenetic relationships among
  the organisms you are studying -- Use different available programs to infer molecular
  evolution, trees and times of divergence.
\item
  Discuss the inferred relationships in the context of previous hypothesis, the
  biology and biogeography of the organisms, etc. -- Answer the question, \emph{is this phylogenetic solution fair/reasonable?}
\end{enumerate}

@@

In the past decade, various studies have developed solutions to automatize the first and second steps, by creating
pipelines that mine already available molecular data from the GenBank repository (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000}),
to obtain homologous characters that can be used for phylogenetic reconstruction.
These tools have been presented as aid for the nonspecialist to decrease some of
the difficulties in the generation of phylogenetic knowledge. However, they are
not that often used as so, suggesting that there are still difficulties for the nonspecialist.
The phylogenetic community has some reserves towards these tools, too. Mainly because they sometimes act as a black box.
However, automatizing the assembly of the character data set is a crucial step towards
reproducibility for a task that was otherwise primarily artisanal and hence largely
non-reproducible.

@@

Public phylogenies can be updated with the ever increasing amount of genetic data that is available on GenBank (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000}).

@@

The goal of Physcraper is to build upon previous phylogenetic knowledge,
allowing a direct comparison between existing phylogenies and phylogenies that are constructed
using new genetic data retrieved from a public nucleotide database (i.e., GenBank (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000})).

@@

\begin{itemize}
\tightlist
\item
  The study tree is a published phylogenetic tree stored in the OToL database, phylesystem (McTavish \emph{et al.} \protect\hyperlink{ref-mctavish2015phylesystem}{2015}). The main
  reason for this is that trees in phylesystem have a set of user defined characteristics
  that are essential for automatizing the phylogeny update process. The most relevant of these being the definition of ingroup and outgroup. Outgroup and ingroup taxa in the original tree are identified and tagged. This allows to automatically set the root for the updated tree on the next steps of the pipeline.
  A user can choose from the `r rotl::tol\_about()\$num\_source\_trees' published trees supporting the resolved node of the synthetic tree in the OToL website (\textless{}\textgreater{}). If the tree you are interested in updating is not in there, you can upload it via OToL's curator tool (\textless{}\url{https://tree.opentreeoflife.org/curator}).
\item
  The alignment should be a gene alignment that was used to generate the tree. The original
  alignments are usually stored in a public repository such as TreeBase (Piel \emph{et al.} \protect\hyperlink{ref-piel2009treebase}{2009}; Vos \emph{et al.} \protect\hyperlink{ref-vos2012nexml}{2012}),
  DRYAD (\url{http://datadryad.org/}), or the journal were the tree was originally published.
  If the alignment is stored in TreeBase, \texttt{physcraper} can download it directly,
  either from the TreeBASE website (\url{https://treebase.org/})
  or through the TreeBASE GitHub repository (SuperTreeBASE; \url{https://github.com/TreeBASE/supertreebase}).
  If the alignment is on another repository, or provided personally by the owner, a copy of it has to be
  downloaded by the user, and it's local path has to be provided as an argument.
\item
  A taxon name matching step is performed to verify that all taxon names on the tips
  of the tree are in the DNA character matrix and vice versa.
\item
  A ``.csv'' file with the summary of taxon name matching is produced for the user.
\item
  Unmatched taxon names are dropped from both the tree and alignment.
  Technically, just one matching name is needed to perform the searches. Please, see next section.
\item
  A ``.tre'' file and a ``.fas'' file containing only the matched taxa are generated and saved in the \texttt{inputs} folder to be used in the following steps.
\end{itemize}

@@

  Users can provide a search taxon that is either a more or a less inclusive
  clade relative to the ingroup of the original phylogeny. If the search taxon is more inclusive, the sequence search will be performed outside the MRCAT of the matched taxa, e.g., including all taxa within
  the family or the order that the ingroup belongs to. If the search taxon is a less inclusive clade, the users can focus on enriching a particular clade/region within the ingroup of the phylogeny.
\item
  The Basic Local Alignment Search Tool, BLAST {[}Altschul \emph{et al.} (\protect\hyperlink{ref-altschul1990basic}{1990}); altschul1997gapped{]} is used to identify
  similarity between DNA sequences within the search taxon in a nucleotide
  database, and the remaining sequences on the alignment.
  The BLAST command line tools (Camacho \emph{et al.} \protect\hyperlink{ref-camacho2009blast}{2009}) are used for both web-based and local-database searches.
\item
  A pairwise alignment-against-all BLAST search is performed. This means that each sequence
  in the alignment is BLASTed against DNA sequences in a nucleotide database constrained to the search
  taxon. Results from each one of these BLAST runs are recorded, and matched sequences are saved
  along with their corresponding identification numbers (accesion numbers in the case of the GenBank database). This information will be used later to store the whole sequences in a dedicated library within the physcraper folder, allowing for secondary analyses to run significantly faster.
\item
  The DNA sequence similarity search can be done on a local database that is easily
  setup by the user. In this case, the BLASTn algorithm is used to performs the similarity search.
\item
  The search can also be performed remotely, on the NCBI database. In this case, the
  bioPython BLAST algorithm is used to perform the similarity search.
\item
  Matched sequences below an e-value, percentage similarity, and outside a minimum
  and maximum length threshold are discarded. This filtering leaves out genomic sequences.
  All acepted sequences are asigned an internal identifier, and are further filtered.
\item
  Because the original alignments usually lack database id numbers, a filtering
  step is needed. Accepted sequences that belong to the
  same taxon of the query sequence, and that are either identical or shorter than
  the original sequence are discarded. Only longer sequences belonging to the
  same taxon as the orignal sequence will be considered further for analysis.
\item
  Among the remaining filtered sequences, there are usually several exemplars per taxon.
  Although it can be useful to keep some of them to, for example, investigate monophyly
  within species, there can be hundreds of exemplar sequences per taxon for some markers.
  To control the number of sequences per taxon in downstream analyses,
  5 sequences per taxon are chosen at random. This number is set by default but can be modified by the user.
\item
  Reverse complement sequences are identified and translated.
\item
  Users can choose to perform a more ``cycles'' of sequence similarity search, by blasting the newly found sequences. This can be done iteratively, but by default only sequences in the alignment are blasted. \textbf{\emph{Is there an argument to control the number of cycles of blast searches with new sequences?}}
\item
  Accepted sequences are downloaded in full, and stored as a local database in a directory that is globally accesible (physcraper/taxonomy), so they are accesible for further runs.
\item
  A fasta file containing all filtered and processed sequences resulting from the BLAST search is generated for the user.
\end{itemize}

@@

\begin{itemize}
\tightlist
\item
  A gene tree is reconstructed for each alignment provided, using the software RAxML (Stamatakis \protect\hyperlink{ref-stamatakis2014raxml}{2014})
  with 100 classic bootstrap (Felsenstein \protect\hyperlink{ref-felsenstein1985confidence}{1985}) replicates by default. The number of bootsrap replicates can be modified by the user.
  Other type of bootstrap that I think is not yet incorporated into physcraper is the Transfer Bootstrap Expectation (TBE) recently proposed in Lemoine \emph{et al.} (\protect\hyperlink{ref-lemoine2018renewing}{2018}).
\item
  The final result is an updated phylogenetic hypothesis for each of the genes provided in the alignment.
\item
  Tips on all trees generated by physcraper are defined by a taxon name space, allowing to perform comparisons and conflict analyses.
\item
  Robinson Foulds metrics
\item
  Describe what a conflict analysis is: Node by node comparison of the resulting clades compared to
\item
  For the conflict analysis to be meaningful, the root of the tree ineeds to be accurately defined.
\item
  Currently, the root is determined by finding the parent node of the sequences that do not belong to the ingroup/ search taxon. This ensures a correct rooting of the tree even when the search taxon is more inclusive than the ingroup.
\item
  Conflict information can only be generated in the context of the whole Open Tree of
  Life. Otherwise, it is not really possible to get conflict data.
  \textbf{\emph{- One way to compare two independent phylogenetic trees is to compare them both to
  the synthetic OToL and then measure how well they do against each other}}
\end{itemize}

@@

The genus \emph{Ilex} is the only extant clade within the family Aquifoliaceae, order Aquifoliales of flowering plants.
It encompasses between 400-600 living species. A review of litterature shows that there are three published phylogenetic trees, showing relationships within the hollies.
The first one has been made available both on OToL phylesystem and synth tree, and on treeBASE, it samples 48 species.
The second has not been made available anywhere, not even in supplementary data of the journal.
\textbf{\emph{Contact authors? They seem old school, probably do not wanna share their data.}}
The most recent one has been made available in the OToL Phylesystem and DRYAD. It is the best sampled yet, with 200 species. However,
it has not been added to the syntehtic tree yet.
This makes it a perfect case to test the basic functionalities of physcraper: we know that the sequences of the most recently published tree have been made available on the GenBank database (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000}). Updating the oldest tree, we should get something very similar to the newest tree.

@@

Let's be more specific now about our X group and say it is the Ascomycota.
The best tree currently available in OToL was published by Schoch \emph{et al.} (\protect\hyperlink{ref-schoch2009ascomycota}{2009}).
The first step, is to get the Open Tree of Life study id. There are some options to do this:
- You can go to the Open Tree of Life website and browse until you find it, or
- you can get the study id using R tools:
- By using the TreeBase ID of the study (which is not fully exposed on the
TreeBase website home page of the study, so you have to really look it up manually):

@@

Phylogeny of the Testudines 6 tips from Crawford \emph{et al.} (\protect\hyperlink{ref-crawford2012more}{2012})
There is just one tree in OToL.
There is just one alignment on \href{https://treebase.org/treebase-web/search/study/matrices.html?id=12742}{treebase} with all the 1 145 loci.

@@

Describe again statistics to compare phylogenies provided by physcraper via OpenTreeOfLife.
Mention statistics provided by other tools: PhyloExplorer (Ranwez \emph{et al.} \protect\hyperlink{ref-ranwez2009phyloexplorer}{2009}).
Compare and discuss.

@@

How can it be used for the advantage of the field:
- rapid phylogenetic placing of newly discovered species, as mentioned in Webb \emph{et al.} (\protect\hyperlink{ref-webb2010biodiversity}{2010})
- obtain trees for ecophylogenetic studies, as mentioned in Helmus \& Ives (\protect\hyperlink{ref-helmus2012phylogenetic}{2012})
- one day could be used to sistematize nucleotide databases, such as Genbank (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000}), as mentioned in San Mauro \& Agorreta (\protect\hyperlink{ref-san2010molecular}{2010}), i.e., curate ncbi taxonomic assignations.
- allows to generate custom species trees for downstream analyses, as mentioned in Stoltzfus \emph{et al.} (\protect\hyperlink{ref-stoltzfus2013phylotastic}{2013})

Things that physcraper does not do:
- analyse the whole GenBank database (Benson \emph{et al.} \protect\hyperlink{ref-benson2000genbank}{2000}; Wheeler \emph{et al.} \protect\hyperlink{ref-wheeler2000database}{2000}) to find homolog regions suitable to reconstruct phylogenies, as mentioned in Antonelli \emph{et al.} (\protect\hyperlink{ref-antonelli2017toward}{2017}). There are already some very good tools that do that.
- provide basic statistics on data availability to assemble molecular datasets, as mentioned by Ranwez \emph{et al.} (\protect\hyperlink{ref-ranwez2009phyloexplorer}{2009}). Phyloexplorer does this?
- it is not a tree repo, as phylota is, mentioned in Deepak \emph{et al.} (\protect\hyperlink{ref-deepak2014evominer}{2014})

@@

Sanderson \emph{et al.} (\protect\hyperlink{ref-sanderson2008phylota}{2008})\strut

@@

Wu \& Eisen (\protect\hyperlink{ref-wu2008simple}{2008})\strut

@@

Smith \emph{et al.} (\protect\hyperlink{ref-smith2009mega}{2009})\strut

@@

Peters \emph{et al.} (\protect\hyperlink{ref-peters2011taming}{2011})\strut

@@

Grant \& Katz (\protect\hyperlink{ref-grant2014building}{2014})\strut

@@

Chesters \& Zhu (\protect\hyperlink{ref-chesters2014protocol}{2014})\strut

@@

Izquierdo-Carrasco \emph{et al.} (\protect\hyperlink{ref-izquierdo2014pumper}{2014})\strut

@@

Meng \emph{et al.} (\protect\hyperlink{ref-meng2015darwintree}{2015}\protect\hyperlink{ref-meng2015darwintree}{a})\strut

@@

Xu \emph{et al.} (\protect\hyperlink{ref-xu2015ncbiminer}{2015})\strut

@@

McMahon \emph{et al.} (\protect\hyperlink{ref-mcmahon2015stbase}{2015})\strut

@@

Papadopoulou \emph{et al.} (\protect\hyperlink{ref-papadopoulou2015automated}{2015})\strut

@@

Kumar \emph{et al.} (\protect\hyperlink{ref-kumar2015bir}{2015})\strut

@@

Antonelli \emph{et al.} (\protect\hyperlink{ref-antonelli2017toward}{2017})\strut

@@

Drori \emph{et al.} (\protect\hyperlink{ref-drori2018onetwotree}{2018})\strut

@@

Smith \& Walker (\protect\hyperlink{ref-smith2019pyphlawd}{2019})\strut

@@

Cerón-Romero \emph{et al.} (\protect\hyperlink{ref-ceron2019phylotol}{2019})\strut

@@

Bennett \emph{et al.} (\protect\hyperlink{ref-bennett2018phylotar}{2018})\strut

@@

According to Cerón-Romero \emph{et al.} (\protect\hyperlink{ref-ceron2019phylotol}{2019}), PhyLoTA and BIR ``focus on the identification and collection
of homologous and paralog genes from public databases such as GenBank'', while both AMPHORA and PHLAWD
``focus on the construction and refinement of robust alignments rather than the collection of homologs.''

@@

PhyloFinder (Chen \emph{et al.} \protect\hyperlink{ref-chen2008phylofinder}{2008}) - cited by 18: a search engine for phylogenetic databases, using
trees from TreeBASE - more related to phylotastic's goal than to updating/creating phylogenies

@@

PhyloExplorer (Ranwez \emph{et al.} \protect\hyperlink{ref-ranwez2009phyloexplorer}{2009}) - cited by 21: a python and MySQL based website to facilitate
assessment and management of phylogenetic tree collections. It provides ``statistics describing the collection,
correcting invalid taxon names, extracting taxonomically relevant parts of the collection
using a dedicated query language, and identifying related trees in the TreeBASE database''.

@@

PhySpeTre (Fang \emph{et al.} \protect\hyperlink{ref-fang2019physpetree}{2019}) - no citations yet - no sequence retrieval, just phylogenetic reconstruction
pipeline.

@@

Datataxa Ruiz-Sanchez \emph{et al.} (\protect\hyperlink{ref-ruiz2019datataxa}{2019}) - no citations yet - focus on extracting metadata from GenBank sequence information.

@@

Phylota was published as a website to summarize and browse the phylogenetic potential of the GenBank
database (Sanderson \emph{et al.} \protect\hyperlink{ref-sanderson2008phylota}{2008}).

@@

    original publication of PHLAWD (Smith \emph{et al.} \protect\hyperlink{ref-smith2009mega}{2009})

@@

    the red algal tree of life (Verbruggen \emph{et al.} \protect\hyperlink{ref-verbruggen2010data}{2010})

@@

    Beaulieu \emph{et al.} (\protect\hyperlink{ref-beaulieu2012modeling}{2012}\protect\hyperlink{ref-beaulieu2012modeling}{a}) cites phylota as an example study of very large and comprehensive

@@

    a review for ecologists about phylogenetic tools (Roquet \emph{et al.} \protect\hyperlink{ref-roquet2013building}{2013})

@@

    a study constructing a dated seed plant phylogeny using pyPHLAWD (Smith \& Brown \protect\hyperlink{ref-smith2018constructing}{2018})

@@

    using genomic data. It aims to be incorporated into a pipeline such as phylota some day (Fan \emph{et al.} \protect\hyperlink{ref-fan2015assembly}{2015}).

@@

    nexml format presentation (Vos \emph{et al.} \protect\hyperlink{ref-vos2012nexml}{2012}) - cites phylota as a tool that uses

@@

    tips of the Campanulidae, constructed with PHLAWD (Beaulieu \& Donoghue \protect\hyperlink{ref-beaulieu2013fruit}{2013})

@@

    a study of Southeast Asia plant biodiversity inventory (Webb \emph{et al.} \protect\hyperlink{ref-webb2010biodiversity}{2010}) -

@@

    a study of wood density for carbon stock assessments (Flores \& Coomes \protect\hyperlink{ref-flores2011estimating}{2011}),

@@

    a study proposing something similar to Open tree but applied only to land plants (Beaulieu \emph{et al.} \protect\hyperlink{ref-beaulieu2012synthesizing}{2012}\protect\hyperlink{ref-beaulieu2012synthesizing}{b})

@@

    an analysis of the phylogenetic diversity-area curve (Helmus \& Ives \protect\hyperlink{ref-helmus2012phylogenetic}{2012}),

@@

    (San Mauro \& Agorreta \protect\hyperlink{ref-san2010molecular}{2010}) cites phylota as a tool ``intended to systematize GenBank information
